---
title: "09 - Shiny"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
You can manipulate the sample size, intercept, slope, and standard deviation of residuals.

2.
The outputs are a graph of OLS best line fit and a table with data on the slope and intercept.

3.
I looked at the Nutritiona Calculator app. In this app, you add data on ingreadients of foods you eat and the app outputs three graphs of macronutrients, minerals and vitamins as % daily value. One cool thing that I found is that nutrient and calorie amounts over 100% are highlighted at the top.

# 9.2 Our First App
```{r}
install.packages("shiny")
library(shiny)


ui <- fluidPage(
  
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)


library(shiny)
ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)
server <- function(input, output, session) {
  # Create a reactive expression
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })

  output$summary <- renderPrint({
    # Use a reactive expression by calling it like a function
    summary(dataset())
  })
  
  output$table <- renderTable({
    dataset()
  })
}
shinyApp(ui, server)
```

```{r}
library(shiny)


ui <- fluidPage(tableOutput("mortgage")
output$greeting <- renderText({
  paste0("Hello ", input$name)
})
numericInput("age", "How old are you?", value = NA)
textInput("name", "What's your name?")
textOutput("greeting")
  
)

server <- function(input, output, session) {output$histogram <- renderPlot({
  hist(rnorm(1000))
}, res = 96)
  
}

shinyApp(ui, server)

```

```{r}
library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", label = "and y is", min = 1, max = 50, value = 30),
  "then x times y is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * y
  })
}

shinyApp(ui, server)
```

```{r}
library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", label = "and y is", min = 1, max = 50, value = 30),
  "then x times y is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * y
  })
}

shinyApp(ui, server)

```

# 9.4 Tennis App
```{r}
atp_df <- read_csv("data/atp_matches_2019.csv")

wta_df <- read_csv("data/wta_matches_2019.csv")

both_df <- bind_rows(atp_df, wta_df)

both_long <- both_df %>% pivot_longer(c(winner_name, loser_name))

## only keep players who have player over 50 matches
both_n50 <- both_long %>% group_by(value) %>% count() %>%
  filter(n > 50)

## construct various statistics
major_tennis <- semi_join(both_long, both_n50, by = c("value"))
major_tennis <- major_tennis %>% mutate(w_svperc = 100 * w_1stIn / w_svpt,
                        l_svperc = 100 * l_1stIn / l_svpt,
                        w_firstwon = 100 * w_1stWon / w_1stIn,
                        l_firstwon = 100 * l_1stWon / l_1stIn,
                        w_secondwon = 100 * w_2ndWon / (w_svpt - w_1stIn),
                        l_secondwon = 100 * l_2ndWon / (l_svpt - l_1stIn))

major_tennis_w <- major_tennis %>% filter(name == "winner_name")
major_tennis_l <- major_tennis %>% filter(name == "loser_name")

w_small <- major_tennis_w %>% select(value, winner_seed, w_ace, w_df, w_svperc,
                                     w_firstwon, w_secondwon) %>%
  rename(seed = winner_seed, ace = w_ace, df = w_df, svperc = w_svperc,
         firstwon = w_firstwon, secondwon = w_secondwon)

l_small <- major_tennis_l %>% select(value, loser_seed, l_ace, l_df, l_svperc, l_firstwon, l_secondwon)  %>%
  rename(seed = loser_seed, ace = l_ace, df = l_df, svperc = l_svperc,
         firstwon = l_firstwon, secondwon = l_secondwon)

df <- bind_rows(w_small, l_small) %>%
  rename(player = "value")
df
```

```{r}
df_sub <- df %>% filter(player == "Daniil Medvedev")
ggplot(df_sub, aes(x = ace)) +
  geom_histogram(colour = "black", fill = "white", bins = 15)

var_choices <- names(df)[3:7]

library(shiny)

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("playerchoice",
                   label = "Choose a Player", 
                   choices = levels(factor(df$player)),
                   selected = "Aryna Sabalenka"),
    radioButtons("varchoice", label = "Choose a Statistic",
                 choices = var_choices),
    sliderInput("binnumber", label = "Choose a Number of Bins", 
                min = 1, max = 50, value = 15, step = 1)),
    mainPanel(plotOutput("histgraph"))
  )
)

server <- function(input, output, session) {
  
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  
  hist_plot <- reactive({
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = input$binnumber) +
      theme_grey(base_size = 22)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  })
  

}

shinyApp(ui, server)
```

# 9.6 Additional Exercises
1.
```{r}
library(shiny)

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("playerchoice",
                   label = "Choose a Player", 
                   choices = levels(factor(df$player)),
                   selected = "Aryna Sabalenka"),
    radioButtons("varchoice", label = "Choose a Statistic",
                 choices = var_choices),
    numericInput("binnumber", label = "Choose a Number of Bins", 
                min = 1, max = 50, value = 15, step = 1)),
    mainPanel(plotOutput("histgraph"))
  )
)

server <- function(input, output, session) {
  
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  
  hist_plot <- reactive({
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = input$binnumber) +
      theme_grey(base_size = 22)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  })
  
  
}

shinyApp(ui, server)
```

2.
```{r}
library(shiny)

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("playerchoice",
                   label = "Choose a Player", 
                   choices = levels(factor(df$player)),
                   selected = "Aryna Sabalenka"),
    radioButtons("varchoice", label = "Choose a Statistic",
                 choices = var_choices),
    sliderInput("binnumber", label = "Choose a Number of Bins", 
                 min = 0, max = 100, value = 5, step = 5, animate =
                  animationOptions(interval = 300, loop = TRUE))),
    mainPanel(plotOutput("histgraph"))
  )
)
server <- function(input, output, session) {
  
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  
  hist_plot <- reactive({
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = input$binnumber) +
      theme_grey(base_size = 22)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  })
  
  
}

shinyApp(ui, server)
```

3.
```{r}
library(shiny)

df_ayo <- df %>%
  filter(major1 == "MATH" | major1 == "STAT" | major1 == "CS")

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(radioButtons(inputId = "majorchoice",
                                label = "Choose a Major",
                                choices = factor(df_ayo$major1))),
    mainPanel(plotOutput(outputId = "majorplot"))
  )
)

server <- function(input, output, session) {
  
  df_update <- reactive({
    df_stat <- df_major %>% filter(discipline == input$majorchoice) 
    
    
    df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
      filter(type == "major1" |
               type == "major2" | 
               type == "major3")
    
    df_nostat <- df_statfull %>% filter(discipline != input$majorchoice &
                                          !is.na(discipline)) %>%
      group_by(discipline) %>%
      summarise(nstudent = n()) %>%
      mutate(discipline = fct_reorder(discipline, nstudent))
  })
  
  
  output$majorplot <- renderPlot({
    ggplot(data = df_update(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })
}

shinyApp(ui, server)
```

4.
```{r}
library(shiny)


ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(selectizeInput(inputId = "majorchoice",
                                label = "Choose a Major",
                                choices = factor(df$major1))),
    mainPanel(plotOutput(outputId = "majorplot"),
              tableOutput(outputId = "majorgender"))
  )
)

server <- function(input, output, session) {
  
  df_update <- reactive({
    df_stat <- df_major %>% filter(discipline == input$majorchoice) 
    
    
    df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
      filter(type == "major1" |
               type == "major2" | 
               type == "major3")
    
    df_nostat <- df_statfull %>% filter(discipline != input$majorchoice &
                                          !is.na(discipline)) %>%
      group_by(discipline) %>%
      summarise(nstudent = n()) %>%
      mutate(discipline = fct_reorder(discipline, nstudent))
  })
  
  df_update2 <- reactive({
    df_stat <- df_major %>% filter(discipline == input$majorchoice)
  })
  
  
  output$majorplot <- renderPlot({
    ggplot(data = df_update(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })
  
  output$majorgender <- renderTable({
    group_by(sex) %>%
      summarise(gender_count = n())
    
  })
}

shinyApp(ui, server)


```
