---
title: "06 - Mapping"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## install.packages("maps")
library(maps)
library(tidyverse)
state_df <- ggplot2::map_data("state")
state_df %>% nrow()
ggplot(data = state_df,
            mapping = aes(x = long, y = lat,
                          group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(axis.text = element_blank()) +
  theme_void()
## install.packages("usdata")
library(usdata)
state_stats
state_stats <- state_stats %>% mutate(state = str_to_lower(state))
state_full <- left_join(state_df, state_stats, by = c("region" = "state"))
state_full <- state_full %>% as_tibble() %>%
  mutate(pop_change = 100 * (pop2010 - pop2000) / pop2000)
ggplot(data = state_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = pop_change)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  scale_fill_distiller(type = "div", direction = 1, limits = c(-1, 1))
## should change so that middle colour is 0
## fermenter is for binning continuous
## brewer is for discrete colours
## distiller is for continuous variable
```

# 6.2
## Exercises
1. Populations of migratory birds over time

2. 

# 6.3 Mapping Other Areas: sf Objects
```{r}
active <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_active_cases.csv", n_max = 34)
tcases <- read_csv("https://raw.githubusercontent.com/iramler/stlawu_covid/main/slc_towns_total_cases.csv", n_max = 34)
active_long <- active %>% pivot_longer(5:ncol(active), names_to = "date",
                                       values_to = "cases")
## repeat for total cases
tcases_long <- tcases %>% pivot_longer(5:ncol(tcases), names_to = "date",
                                       values_to = "cases")
library(lubridate)
covid_df <- left_join(tcases_long, active_long,
                      by = c("date", "Order", "NAME")) %>%
  mutate(date = mdy(date)) %>%
  rename(total_cases = cases.x,
         active_cases = cases.y) %>%
  mutate(total_cases = if_else(is.na(total_cases),
                               true = 0, false = total_cases),
         active_cases = if_else(is.na(active_cases),
                                      true = 0, false = active_cases))
  
covid_SLC <- covid_df %>% filter(NAME == "St. Lawrence County")
covid_sub <- covid_df %>% filter(NAME != "St. Lawrence County")
covid_SLC
covid_sub
```

## Exercises
1.
```{r}
ggplot(data = covid_SLC, aes(x = date, y =active_cases)) +
  geom_line()
## install.packages("sf")
library(sf)
shp <- read_sf("data/SLC_Civil_Boundaries_SHP/slc.shp") %>%
  st_transform(st_crs("+proj=longlat"))
shp
ggplot(data = shp) +
  geom_sf() +
  theme_void()
full_df <- left_join(shp, covid_sub, by = "NAME") %>%
  filter(date == max(date)) ## only plot cases on the most recent date
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void()
```

2.
```{r}
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() +
  scale_colour_brewer(palette = "Blues")
```
Sequential

3.
```{r}
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() +
  scale_fill_viridis_b()
```

4.
```{r}
ggplot(data = full_df) +
  geom_sf(aes(fill = active_cases)) +
  theme_void() +
  scale_fill_viridis_b() +
  geom_sf_text(aes(label = active_cases))
```

# 6.4 Leaflet
```{r}
install.packages("devtools")
library(devtools)
devtools::install_github("li-wen-li/uszipcodes")
library(uszipcodes)
beers <- read_csv("data/breweries.csv") 
raw_zip <- uszipcodes::get_zip(beers$address)
beers$Zip <- as.integer(uszipcodes::clean_zip(raw_zip))
## only keep zip, lat, and long
zip_tab <- zip_table %>% dplyr::select(Zip, Latitude, Longitude)
beer_location <- inner_join(beers, zip_tab)
content <- beer_location %>%
  mutate(popup = paste0('<a href =', beer_location$website, '>',
                        beer_location$brewery_name, '</a>'))
install.packages("leaflet")
library(leaflet)
beer_map <- leaflet(beer_location) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 4) %>% 
  addTiles() %>% 
  addProviderTiles(providers$Stamen.Watercolor) %>% 
  addMarkers(lng = beer_location$Longitude, lat = beer_location$Latitude,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
beer_map
```

## Exercises
1. It gets rid of breweries with zip codes that do not match while preserving all variables in each df.

# 6.6 Supplementary Exercises

1.
```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")


map <- map_data("world", region = ".", exact = FALSE)

map2 <- map %>%
  mutate(Country = region)

map2 <- inner_join(hpi_df, map2, by = "Country")

ggplot(data = map2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = HPIRank)) +
  coord_map(projection = "mercator") 
```


2.
```{r}
map2 <- map %>%
  mutate(Country = region)

map2$Country[map2$Country=="USA"] <- "United States of America"

map2 <- full_join(hpi_df, map2, by = "Country")

library(gridExtra)
ggplot(data = map2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = HPIRank)) +
  coord_map(projection = "mercator")
```

3.
```{r}
ggplot(data = map2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = HPIRank)) +
  coord_map(projection = "mercator", xlim=c(-180,180))

```

4.
Greenland can actually fit inside Africa 14 times over.

5.
```{r}
ggplot(data = map2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = HPIRank)) +
  coord_map(projection = "globular", xlim=c(-180,180))

ggplot(data = map2, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = HPIRank)) +
  coord_map(projection = "gilbert", xlim=c(-180,180))

```

The size comparison is more realistic in both, but the globular projection seems to be the most accurate out of the three.